Improving real-time trend estimates using local parametrisation of local polynomial regression filters

### 2024-02-12 article_jos.qmd
transpose vector in inline text

Let $p$ et $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\begin{pmatrix}\theta_{-p}\\\vdots\\\theta_{f}\end{pmatrix}$ such as for all time series $X_t$:
$$
M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
$$

@@

To estimate $\boldsymbol \beta$ we need $2h+1\geq d+1$ and the estimation is made by weighted least squares (WLS), which is equivalent to minimising the following objective function:
$$
S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
$$
where $\kappa_j$ is a set of weights called *kernel*.
We have $\kappa_j\geq 0:\kappa_{-j}=\kappa_j$, and with $\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})$, the estimate of $\boldsymbol \beta$ can be written as $\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.$
With $\boldsymbol e_1=\begin{pmatrix}1 \\0 \\\vdots\\0 \end{pmatrix}$, the estimate of the trend is:
$$
\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\boldsymbol \theta'\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
$$ {#eq-mmsym}
To conclude, the estimate of the trend $\hat{m}_{t}$ can be obtained applying the symmetric filter $\boldsymbol \theta$ to $y_t$ ($\boldsymbol \theta$ is symmetric due to the symmetry of the kernel weights $\kappa_j$).
Moreover, $\boldsymbol X'\boldsymbol \theta=\boldsymbol e_{1}$ so:
$$
\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\llbracket 1,d\right\rrbracket:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
$$
Hence, the filter $\boldsymbol w$ preserve deterministic polynomial of order $d$.

@@

The initial series is modelled here as:
$$
\boldsymbol y=\boldsymbol U \boldsymbol \gamma+\boldsymbol Z\boldsymbol \delta+\boldsymbol \varepsilon,\quad
\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
$$ {#eq-lpgeneralmodel}
where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X$.
The objective is to find a filter $\boldsymbol v$ that minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain constraints.
These constraints are represented by the matrix $\boldsymbol U=\begin{pmatrix}\boldsymbol U_{p}\\\boldsymbol U_{f}\end{pmatrix}$: $\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta$ (with $\boldsymbol U_p$ the matrix $(h+q+1)\times (d+1)$ which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
The problem is equivalent to finding $v$ which minimises:
$$
\varphi(\boldsymbol v)=
\underbrace{
  \underbrace{(\boldsymbol v-\boldsymbol \theta_{p})'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
  \boldsymbol \theta_{f}'\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
  \underbrace{[\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}}_{bias^2}
}_\text{mean square revision error}+
\underbrace{2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)}_{\text{constraints}}.
$$ {#eq-lppasym}
where $\boldsymbol l$ is a vector of Lagrange multipliers.

@@

When $\boldsymbol U=\begin{pmatrix}1\\\vdots\\1\end{pmatrix}$, $\boldsymbol Z=\begin{pmatrix}-h\\\vdots\\+h\end{pmatrix}$, $\boldsymbol \delta=\delta_1$, $\boldsymbol D=\sigma^2\boldsymbol I$ and when the symmetric filter is the Henderson filter, we find the asymmetric Musgrave filters (using the Henderson kernel).
This filter assumes that, for real-time estimation, the data is generated by a linear process and that the asymmetric filters preserve the constants ($\sum v_i=\sum \theta_i=1$).
These asymmetric filters depend on the ratio $\lvert\delta_1/\sigma\rvert$, which, assuming that the trend is linear and the bias constant, can be linked to the I-C ratio $R=\frac{\bar{I}}{\bar{C}}=\frac{\sum\lvert I_t-I_{t-1}\rvert}{\sum\lvert C_t-C_{t-1}\rvert}$: $\delta_1/\sigma=2/(R\sqrt{\pi})$.
This ratio is used in particular in X-13ARIMA to determine the length of the Henderson filter. 
For monthly data:

@@

Using the same notation as in @sec-lppasymf, $\boldsymbol\theta$ represents the symmetric filter and $\boldsymbol v$ the asymmetric filter.
With $\boldsymbol\theta=\begin{pmatrix}\boldsymbol\theta_p\\\boldsymbol\theta_f\end{pmatrix}$ and $\boldsymbol\theta_p$ of the same length of $\boldsymbol v$, and $\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p$, the timeliness criterion can be expressed as:
$$
T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
\quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
$$
Furthermore, the objective function $\varphi$ of @eq-lppasym can be rewritten:
\begin{align*}
\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
\begin{cases}
\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
\boldsymbol c\text{ a constant independent of }\boldsymbol v
\end{cases}.
\end{align*}
Adding the timeliness criterion:
$$
\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
\widetilde {\boldsymbol c}\quad\text{with }
\begin{cases}
\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
\end{cases},
$$
where $\alpha_T$ is the weight associated to the timeliness criterion. 
With $\alpha_T=0$ we find $\varphi(\boldsymbol v)$.
This extension therefore makes it possible to find all the symmetric and asymmetric filters presented in the previous section but also generalizes the approach of @GrayThomson1996.
This extension is implemented in the R function `rjd3filters::lp_filter()`.

>>>>

Let $p$ et $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $X_t$:
$$
M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
$$

@@

To estimate $\boldsymbol \beta$ we need $2h+1\geq d+1$ and the estimation is made by weighted least squares (WLS), which is equivalent to minimising the following objective function:
$$
S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
$$
where $\kappa_j$ is a set of weights called *kernel*.
We have $\kappa_j\geq 0:\kappa_{-j}=\kappa_j$, and with $\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})$, the estimate of $\boldsymbol \beta$ can be written as $\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.$
With $\boldsymbol e_1=\transp{\begin{pmatrix}1 &0 &\cdots&0 \end{pmatrix}}$, the estimate of the trend is:
$$
\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\boldsymbol \theta'\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
$$ {#eq-mmsym}
To conclude, the estimate of the trend $\hat{m}_{t}$ can be obtained applying the symmetric filter $\boldsymbol \theta$ to $y_t$ ($\boldsymbol \theta$ is symmetric due to the symmetry of the kernel weights $\kappa_j$).
Moreover, $\boldsymbol X'\boldsymbol \theta=\boldsymbol e_{1}$ so:
$$
\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\llbracket 1,d\right\rrbracket:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
$$
Hence, the filter $\boldsymbol w$ preserve deterministic polynomial of order $d$.

@@

The initial series is modelled here as:
$$
\boldsymbol y=\boldsymbol U \boldsymbol \gamma+\boldsymbol Z\boldsymbol \delta+\boldsymbol \varepsilon,\quad
\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
$$ {#eq-lpgeneralmodel}
where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X$.
The objective is to find a filter $\boldsymbol v$ that minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain constraints.
These constraints are represented by the matrix $\boldsymbol U=\transp{\begin{pmatrix} \transp{\boldsymbol U_{p}}& \transp{\boldsymbol U_{f}}\end{pmatrix}}$: $\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta$ (with $\boldsymbol U_p$ the matrix $(h+q+1)\times (d+1)$ which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
The problem is equivalent to finding $v$ which minimises:
$$
\varphi(\boldsymbol v)=
\underbrace{
  \underbrace{(\boldsymbol v-\boldsymbol \theta_{p})'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
  \boldsymbol \theta_{f}'\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
  \underbrace{[\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}}_{bias^2}
}_\text{mean square revision error}+
\underbrace{2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)}_{\text{constraints}}.
$$ {#eq-lppasym}
where $\boldsymbol l$ is a vector of Lagrange multipliers.

@@

When $\boldsymbol U=\transp{\begin{pmatrix}1&\cdots&1\end{pmatrix}}$, $\boldsymbol Z=\transp{\begin{pmatrix}-h&\cdots&+h\end{pmatrix}}$, $\boldsymbol \delta=\delta_1$, $\boldsymbol D=\sigma^2\boldsymbol I$ and when the symmetric filter is the Henderson filter, we find the asymmetric Musgrave filters (using the Henderson kernel).
This filter assumes that, for real-time estimation, the data is generated by a linear process and that the asymmetric filters preserve the constants ($\sum v_i=\sum \theta_i=1$).
These asymmetric filters depend on the ratio $\lvert\delta_1/\sigma\rvert$, which, assuming that the trend is linear and the bias constant, can be linked to the I-C ratio $R=\frac{\bar{I}}{\bar{C}}=\frac{\sum\lvert I_t-I_{t-1}\rvert}{\sum\lvert C_t-C_{t-1}\rvert}$: $\delta_1/\sigma=2/(R\sqrt{\pi})$.
This ratio is used in particular in X-13ARIMA to determine the length of the Henderson filter. 
For monthly data:

@@

Using the same notation as in @sec-lppasymf, $\boldsymbol\theta$ represents the symmetric filter and $\boldsymbol v$ the asymmetric filter.
With $\boldsymbol\theta=\transp{\begin{pmatrix}\transp{\boldsymbol\theta_p}&\transp{\boldsymbol\theta_f}\end{pmatrix}}$ and $\boldsymbol\theta_p$ of the same length of $\boldsymbol v$, and $\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p$, the timeliness criterion can be expressed as:
$$
T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
\quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
$$
Furthermore, the objective function $\varphi$ of @eq-lppasym can be rewritten:
\begin{align*}
\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
\begin{cases}
\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
\boldsymbol c\text{ a constant independent of }\boldsymbol v
\end{cases}.
\end{align*}
Adding the timeliness criterion:
$$
\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
\widetilde {\boldsymbol c}\quad\text{with }
\begin{cases}
\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
\end{cases},
$$
where $\alpha_T$ is the weight associated to the timeliness criterion. 
With $\alpha_T=0$ we find $\varphi(\boldsymbol v)$.
This extension therefore makes it possible to find all the symmetric and asymmetric filters presented in the previous section but also generalizes the approach of @GrayThomson1996.
This extension is implemented in the R function `rjd3filters::lp_filter()`.
